### 2020-04-29 final_report/results.tex
revise scaling benchmark results

@@

In the next experiment we examine surrogate scaling properties correlating
metrics of interest with progressively increasing training set size. The results
shown in~\cref{fig:scaling} seem to suggest that in terms of regression performance,
methods based on decision trees and artificial neural networks offer the best regression performance on
larger sets overall. This is a curious extension of the previous findings, where
gradient boosted trees were observed to significantly dominate over the rest of
the examined methods. With increasing training set size, their relative advantage is
clearly gradually diminished.

\begin{figure}[h]
	\centering
	\begin{subfigure}[b]{0.333\textwidth}
		\centering
		\includegraphics[width=\linewidth]{scaling_metric_r2}
		% TODO: only a placeholder, regenerate when data becomes available
		\caption{Regression performance (as $R^2$)}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.333\textwidth}
		\centering
		\includegraphics[width=\linewidth]{scaling_time_train}
		% TODO: only a placeholder, regenerate when data becomes available
		\caption{Complexity (as~$\overline{t}_{\text{trn.}}$)}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.333\textwidth}
		\centering
		\includegraphics[width=\linewidth]{scaling_time_pred}
		% TODO: only a placeholder, regenerate when data becomes available
		\caption{Complexity (as~$\overline{t}_{\text{pred.}}$)}
	\end{subfigure}
	\caption{Various metrics collected during experiment 3 (scaling
	benchmark) displayed as a function of training set size.}
	\label{fig:scaling}
\end{figure}

According to our experiment, the lowest mean training time is generally achieved
by instance-based learning methods, which seem to offer near-constant scaling
characteristics at the expense of significant performance increase later during
prediction. Following that, we observe that the majority of tree-based methods also exhibit
desirable properties. The notable exception here appear to be artificial neural networks,
which are the only model to utilise parallelisation. As such, their constant
synchronisation overhead presumably hinders performance on small training sets,
producing misleading results when divided by the number of samples.

In terms of mean prediction time, all tested surrogates except previously mentioned
instance-based learning methods scale exceptionally well. Tree-based generally
models appear to perform the fastest.

>>>>

Following both hyperparameter tuning experiments, we conclude that while domain
restrictions employed in the first case have proven effective in improving the
regression performance of some methods, this result has fluctuated considerably
depending on the selected slices. Furthermore, in all instances the best
results were achieved by families of surrogates that were nearly unaffected by
this modification.

@@

In the third experiment we examine surrogate scaling properties by correlating
metrics of interest with training set size. First, the results shown 
in~\cref{fig:scaling-r2} suggest that the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training data. While such families achieve nearly comparable regression
performance on the largest dataset, in the opposite case methods based on
decision trees clearly outperform neural networks. This can be observed
particularly on training sets of sizes up to~\num{6000}.

\begin{figure}[h]
	\centering
	\begin{subfigure}[b]{0.333\textwidth}
		\centering
		\includegraphics[width=\linewidth]{scaling_metric_r2}
		\caption{Regression performance (as $R^2$)}
		\label{fig:scaling-r2}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.333\textwidth}
		\centering
		\includegraphics[width=\linewidth]{scaling_time_train}
		\caption{Complexity (as~$\overline{t}_{\text{trn.}}$)}
		\label{fig:scaling-trn}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.333\textwidth}
		\centering
		\includegraphics[width=\linewidth]{scaling_time_pred}
		\caption{Complexity (as~$\overline{t}_{\text{pred.}}$)}
		\label{fig:scaling-pred}
	\end{subfigure}
	\caption{Various metrics collected during experiment 3 (scaling
	benchmark) displayed as a function of training set size.}
	\label{fig:scaling}
\end{figure}

Next, we examine scaling behaviour in terms of the mean training time (displayed
in~\cref{fig:scaling-trn}). Consistent with our expectation, the shortest times
were achieved by instance-based learning methods (e.g. KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also exhibit
desirable properties unlike radial basis functions and gaussian process
regression, which appear to scale superlinearly. We note that artificial neural networks,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. Our conjecture is that this effect may be caused
by a constant multi-threading overhead that possibly dominates the training process
on relatively small training sets.

Lastly, we study scaling with respect to the mean prediction time (shown
in~\cref{fig:scaling-pred}). Our initial observation is that all tested
surrogate families with the exception of previously mentioned instance-based
learning methods offer desirable characteristics overall. Analogous to previous
experiments, gradient boosted trees and artificial neural networks again
appear to be tied, as they not only exhibit comparable times but also similar
scaling slopes.
